A Wait-free Queue with Poly-logarithmic Step Complexity

### 2022-12-18 podc_paper/algorithm.tex
started working on description of the algorithm

We use a static {\it tournament tree}, which is a static binary tree of height $\ceil{\log_2 p}$ with one leaf 
assigned to each process.

There is a shared binary tree among the processes to agree on one
total ordering of the operations invoked by processes. The tree is
called a \it{tournament tree} (see Figure~\ref{fig::blocktree}). Each
process has a leaf in which the operations invoked by the process are
stored in order. When a process wishes to do an operation it appends
the operation to its leaf and tries to propagate its new operation up
to the tree's root. Each node of the tree keeps an ordering of
operations propagated up to it. All processes agree on the sequence of
operations in the root and this ordering is used as the linearization
ordering.

Each of the processes $P_1,P_2,...,P_p$ has a leaf and in each node
there is an ordering of operations stored. Each process tries to
propagate its operations up to the root, which stores a total ordering
of all operations.

>>>>

We use a \emph{tournament tree} to agree on a total ordering of the operations performed on the queue.
The tree is a static binary tree of height $\ceil{\log_2 p}$ with one leaf 
assigned to each process. 
Each node of the tree stores an array of \emph{blocks} and each \emph{block} represents a 
sequence of enqueues and a sequence of dequeues.
To perform an operation on the queue, a process $P$ creates a block containing that single 
operation and appends it to the array in $P$'s leaf.
Then, $P$ attempts to propagate the operation along the path from that leaf to the root of the tree.
At each node $v$ along this path, $P$ combines information from the blocks in $v$'s children into a single
block and attempts to append this block to $v$'s array using a CAS instruction.
This cooperative approach is sufficient to ensure that if $P$ fails to CAS a new block into $v$'s array twice,
then its operation has been propagated to $v$ by some other process.
We define a total order on all operations that have been propagated to the root, which is used to
linearize the operations.
In this section, we assume for simplicity that each node has an infinite blocks array.
We describe in Section \ref{??} how to replace the infinite array by a representation that uses bounded space.

If $P$'s operation is an enqueue, it can terminate when the operation has been propagated to the root.
If $P$'s operation is a dequeue, $P$ must use information in the tree to compute the value that the
dequeue must return.  To do this, $P$ first determines which block in the root contains its operation
(since the operation may have been propagated to the root by some other process).
Then, $P$ determines whether the queue is empty when its dequeue is linearized. If so, it returns \nil.
If not, $P$ computes the rank $r$ of its dequeue in the linearization ordering
among all dequeues that do not return \nil,
finds the $r$th enqueue in the linearization and returns that enqueue's value.

The primary challenge is thus figuring out what information to store in each block so that 
the following tasks can be done efficiently (in a polylogarithmic number of steps).
\begin{itemize}
\item
Construct a block for node $v$ that represents the operations contained in $O(p)$ consecutive blocks in $v$'s children.
\item
Given an operation in a leaf that has been propagated to the root, find that operation's location in the root's blocks array.
\item
Given the location of a dequeue operation in the root's blocks array, find the enqueue whose value it should return (or determine that the queue is empty when the dequeue is linearized).
\end{itemize}

---------
